---
title: "Exercise 6"
author: "Carilli"
date: '`r format(Sys.Date(), "%Y-%B-%d")`'
output: 
  html_document:
    theme: cerulean
    highlight: tango
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment = NA, warning = FALSE, message = FALSE,
                      options(digits = 3, scipen = 999))
library(tidyverse)
```

```{r 6data}
globaled <- 
  here::here("data", "globaled.csv") %>% 
  read_csv() %>% 
  janitor::clean_names()
```

### (a)

```{r 6a}
globaled %>% 
  lm(ypcgr ~ testavg + edavg + ypc60, .) %>% 
  jtools::summ()

globaled %>% 
  lm(ypcgr ~ testavg + edavg + ypc60, .) %>%   
  lm.beta::lm.beta()

globaled %>% 
  lm(ypcgr ~ testavg + edavg + ypc60, .) %>% 
  jtools::summ(scale = TRUE, transform.response = TRUE)

globaled %>% 
  mutate(across(where(is.numeric), scale)) %>% 
  lm(ypcgr ~ testavg + edavg + ypc60 -1, .) %>%
  jtools::summ()
```

### (b)

We are testing:

$$\begin{aligned}
H_0&: \beta_1=\beta_2 \\
H_1&: \beta_1\ne\beta_2
\end{aligned}$$

The unrestricted model is

$$\text{UR:}\hspace{1cm}ypcgr=\beta_0+\beta_1testavg+\beta_2edavg+\beta_3ypc60+\epsilon$$

To form the restricted model let $\beta_1=\beta_2$.  This yields:

$$\text{R:}\hspace{1cm}ypcgr=\beta_0+\beta_1(testavg + edavg)+\beta_3ypc60+\nu$$

There are three ways we can test this hypothesis. 

(@) Direct t-test where 

$$t\sim \frac{\left(\hat{\beta_1}-\hat{\beta_2}\right) - \left(\beta_1-\beta_2 \right) }{se_{\left(\hat{\beta_1}-\hat{\beta_2} \right)}}$$
where $$se_{\left(\hat{\beta_1}-\hat{\beta_2} \right)}=\sqrt{\text{Var}\left(\hat{\beta_1}\right) + \text{Var}\left(\hat{\beta_2}\right) - 2\times\text{Cov}\left(\hat{\beta_1},\hat{\beta_2}\right)}$$

```{r 6b_1}
ols_s <- 
globaled %>% 
  mutate(across(where(is.numeric), scale)) %>% 
  lm(ypcgr ~ testavg + edavg + ypc60 -1, .) 

#retrieve the coefficients

beta <- ols_s$coefficients

#generate the Var-Cov matrix 

var_cov <- vcov(ols_s)

df <- ols_s$df.residual 

(t <- (beta[1] - beta[2]) / sqrt(var_cov[1,1] + var_cov[2,2] - 2 * var_cov[1,2]))

qt(.975, df)

(1 - pt(t, df))*2

```

(@) Indirect t-test

Define a new parameter $\delta$ which takes on the value 0 under the null hypothesis. In this case $\delta=\beta_1-\beta_2$. Substitute into the unrestricted equation as $\beta_2=\beta_1 - \delta$, estimate 

$$\begin{aligned}ypcgr&=\beta_0+\beta_1testavg + \left(\beta_1-\delta \right)edavg + \beta_3ypc60+\mu\\
&= \beta_0+\beta_1testavg + \beta_1 edavg - \delta edavg + \beta_3ypc60+\mu\\
&= \beta_0+\beta_1(testavg + edavg) - \delta edavg + \beta_3ypc60+\mu\\
\end{aligned}$$

and test $$H_0:\delta=0 \\ H_1:\delta\ne0$$

```{r 6b_2}
globaled %>% 
  mutate(across(where(is.numeric), scale)) %$%
  lm(ypcgr ~ I(testavg + edavg) + edavg + ypc60 -1, .) %>% 
  jtools::summ()
```

(@) F-test (Wald test)

Where $$F \sim \frac{\frac{ess_R-ess_{UR}}{1}} {\frac{ess_{UR}}{n-k-1}}$$

```{r 6b_3}
ols_r <- 
globaled %>% 
  mutate(across(where(is.numeric), scale)) %>%
  lm(ypcgr ~ I(testavg + edavg) + ypc60 -1, .) %>% 
  jtools::summ()

ols_u <- 
  globaled %>% 
  mutate(across(where(is.numeric), scale)) %>%
  lm(ypcgr ~ testavg + edavg + ypc60 -1, .) %>% 
  jtools::summ()

ess_r <- 
  ols_r$residuals^2 %>% 
  sum()
 
ess_u <- 
  ols_u$residuals^2 %>% 
  sum()
   
(f_c <- (ess_r-ess_u)/(ess_u/df))

qf(.95, 1, df, lower.tail = TRUE)

pf(f_c, 1, df, lower.tail = FALSE)

```

We can perform this test by estimating each model and comparing the appropriate sum of squares or would could perform the test with `linearHhypothesis` from the `car` package as described in the companion.

```{r 6b_4}
ols_s %>% 
  car::linearHypothesis("testavg = edavg")
```

Note the p-values are the same an each case, the calculated t-stats are the same in parts 1 and 2.  Finally, note that the calculate F is the square of the calculate t. This is always the case when there is one degree of freedom in the numerator of an F-test.

### (c)

```{r 6c_i}
# we can also use scale from the base package to standardize the variables
globaled %>% 
  mutate(across(where(is.numeric), scale)) %>%   
  lm(ypcgr ~ testavg + edavg + ypc60 + proprts + open - 1, .) %>% 
  jtools::summ()

```

```{r 6c_ii}
globaled %>% 
  mutate(across(where(is.numeric), scale)) %>% 
  lm(ypcgr ~ testavg + edavg + ypc60 + proprts + open - 1, .) %>% 
  car::linearHypothesis("testavg = proprts")
```



